Report

### 2015-08-06 Report April 2014/draft.tex
Reorganized Report Record

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[12pt,a4paper,final]{article}
%\documentclass[conference]{IEEEtrans}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
\usepackage[latin1]{inputenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
\usepackage{psfrag}
\usepackage{color}
\usepackage{epsfig}
\usepackage{appendix}
\usepackage{caption}
\usepackage{cite}
\usepackage{ifpdf}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{stfloats}
\usepackage{url}
\usepackage{fixltx2e}
\usepackage{setspace} 
\usepackage{multirow}

@@

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Parallel Acceleration of Fixed Complexity Sphere Decoder for Large MIMO Uplink System based on GPU}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{\IEEEauthorblockN{Tianpei Chen}
%\IEEEauthorblockA{ID: 25504946\\Department of Electrical and\\Computer Engineering\\
%McGill University\\
%Montreal, Quebec}}
\author{Tianpei Chen, Harry Leib\\
 Department of Electrical and Computer Engineering\\
 McGill University\\
 Montreal, Quebec}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Large MIMO constitutes a principal techniques for next generation wireless communication systems. Detection algorithms for large MIMO tend to require long simulation times and on conventional computer systems, In order to reduce the simulation times for decoders of large MIMO systems. In this paper we propose to use General Purpose Graphic Processing Units (GPGPU)to implement Fixed Complexity Sphere Decoder (FCSD) for the large scale MIMO system based on the dominant framework of GPGPU-Computed Unified Device Architecture (CUDA). The proposed CUDA-FCSD can chieve up to five times speedup over conventional CPU implementation while reaching the same Bit Error Rate (BER) performance. Furthermore the proposed algorithm demonstrates great potential to accelerate large MIMO systems.
\end{abstract}

@@

\section{Introduction}
 \paragraph{} The present challenge in the wireless communication area is the sharp contrast between the increasing demand of mobile network speed for the transmission of rich media content with better QoS (quality of service), and the shortage of radio frequency spectrum. Multiple Input Multiple Output (MIMO) is thought to be a very promising technology that can help to solve this problem owing to several advantages it can offer. The core idea of MIMO is to use multiple antennas that at the transmitting and receiving sides that provide parallel data pipes in space resulting in spatial multiplexing gains\cite{oestges2010mimo}. For example the LTE standard supports up to 8 antennas at the base station\cite{dahlman20103g}. 
 \paragraph{} Large scale MIMO (also named massive MIMO) systems employ hundreds of low-power low-price antennas collocated at the base station and serving several tens of terminals at the same time\cite{rusek2013scaling}. This technology originated from multiuser MIMO (MU-MIMO) systems\cite{larsson2013massive}. Large scale MIMO will inherit and enhance all the advantages of conventional MIMO (spatial diversity gain and spatial multiplexing gain), as well as reduction of latency and improvement of system robustness. 
 A common decoder for MIMO system is the Sphere Decoder (SD) that perform as well as a Maximum Likelihood (ML) decoder with a dramatic complexity reduction\cite{viterbo1999universal}\cite{hassibi2005sphere}. There are two major problems with the traditional SD decoder: complexity depending on Signal to Noise Ratio (SNR), and sequential nature, which is contradictory with the requirement of a constant data processing throughput and parallel implementation for practical system. The Fixed Complexity Sphere Decoder (FCSD) overcomes these two disadvantages of the SD decoder by searching the solution along a fixed number of independent paths\cite{barbero2008fixing}.
\paragraph{}General purpose Graphic processing Units Computing (GPGPU) has become a common trend in industry and research in areas that need massive computations such as computer vision\cite{fung2008using}, computer games\cite{blewitt2013applicability}, signal processing\cite{van2011accelerating}\cite{6671435} and finance\cite{grauer2013accelerating}. Compared to CPU, the GPU can naturally cope with computational intensive tasks, since most of the hardware resource are allocated for data processing and parallel computation, rather than caching and flow control. In recent years GPU computing has provided significant advantages in data processing speeds over CPU computations\cite{nvidia2008programming}. Computed unified Device Architecture (CUDA) is a programming model for GPGPU that was proposed by Nvidia and is becoming the dominant framework for GPGPU programming. It is based on popular programming languages(C, C++, Java, Python, Fortran, Direct Compute), such that the application developers in different areas do not need to learn a new programming language or get to know the hardware details. CUDA employs the Single Instruction Multiple Threads (SIMT) programming model, enables a GPU implement parallel data processing, and has become a parallel computing platform for wide community of different applications.
\paragraph{}In this paper, a GPU based implementation of Fixed Complexity Sphere Decoder is proposed, A heterogeneous programming solution for FCSD preprocessing and blocked parallel path searching solution are presented, the speedup results of different influence factors are presented and analyzed, The proposed CUDA-FCSD implementation demonstrates a great potential for large MIMO systems.
\paragraph{}The rest of the paper is organized as follows. Section \ref{system} describes the MIMO systems model. Section \ref{programming model} introduces CUDA GPU architecture and the programming model. In section \ref{GPUFCSD}, we present details of the implementation of CUDA-FCSD. In section \ref{simulation}, the speedup and bit error rate (BER) performance of CUDA-FCSD are presented and analyzed. Finally conclusions are drawn in section \ref{conclusion}.    
\section{MIMO System Model}\label{system}
% no \IEEEPARstart
%This demo file is intended to serve as a ``starter file''
%for IEEE conference papers produced under \LaTeX\ using
%IEEEtran.cls version 1.8 and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
%I wish you the best of success.
%\hfill mds
%\hfill December 27, 2012
\paragraph{}We consider a complex uncoded spatial multiplexing MIMO system with $N_r$ receive and $N_t$ transmit antennas, $N_{r}\geq N_{t}$, over a flat fading channel, corresponding to the discrete time model:
\mathbf{y}=\mathbf{H}\mathbf{s}+\mathbf{n}   \label{formula 1}
$\mathbf{s}\in \mathbb{C}^{N_{t}\times 1}$ is the transmitted symbol vector, with components that are mutually independent and taken from the signal constellation $\mathbb{O}$ (4-QAM, 16-QAM, 64-QAM), of size $M$. The transmitted symbol vectors forms a $N_{t}$ complex signal vector space $\mathbb{O}^{N_{t}}$. $\mathbb{E}[\mathbf{s}\mathbf{s}^{H}]=\mathbf{I}_{N_t}E_{s}$, $E_{s}$ denotes the symbol average energy, and $\mathbb{E}[\cdot]$ denotes the expectation operation. Furthermore,  $\mathbf{y}\in \mathbb{C}^{N_{r}\times 1}$ is the received symbol vector, $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ denotes the Rayleigh fading channel propagation matrix with independent identity distributed(i.i.d) circularly symmetric complex Gaussian zero mean components of unit variance, $\mathbf{n}\in \mathbb{C}^{N_{r}\times 1}$ is the additive white Gaussian noise (AWGN) samples with zero mean components and $\mathbb{E}[\mathbf{n}\mathbf{n}^{H}]=\mathbf{I}_{N_{r}}N_{0}$, $N_{0}$ denotes the average noise power spectrum density and $\frac{E_{s}}{N_{0}}$ denotes the signal to noise ratio (SNR). \\

\paragraph{} Assume we have the perfect channel state information (CSI), meaning that $ \mathbf{H}$ is known at the receiver, as well as the SNR. The task of MIMO decoder is to estimate $\mathbf{s}$ based on $\mathbf{y}$ and $\mathbf{H}$.

@@

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

\section{CUDA GPU Architecture and  Programming Model} \label{programming model}
\subsection{Memory Hierarchy}  
CUDA programming model is the C programming extension, with special defined syntax and structure. The serial code executes on the host (CPU) side, and massive parallel data processing is performed on the device (GPU) side, is defined as a heterogeneous programming model. A thread is the basic programming unit of CUDA, a block of cooperative threads form a thread block, in one block all the threads can share one pitch of on chip memory (shared memory), all the threads executed in one block can synchronize their action meaning the threads are permitted to run to the next instruction only when all the threads have reached the same instruction node. Blocks can be grouped into a one-dimensional, two-dimensional, or three-dimensional grid. Similarly, The threads can also be organized into one-dimensional, two-dimensional or three-dimensional blocks. The organization of the block and grid shape and dimension is determined by the application, Figure \ref{figure1} illustrates the 3 dimensional thread hierarchy of CUDA.
\begin{figure}[htb]
\centering
\includegraphics[scale=0.5]{CUDA_programming_hierachy.eps}
\caption{CUDA 3D thread hierachy\\ $5\times 5\times 5$ grid and each block contains $5\times 5\times 5$ threads\\}
\label{figure1}
\end{figure}

\subsection{Single Instruction Multiple Thread (SIMT) Model} Threads are scheduled, managed, created and executed in a group of 32 threads called warp, the programs executes on the device side are defined as $\mathit{kernel}$ function, when a $\mathit{kernel}$ function is invoked, the thread blocks are partitioned into warps, the threads in one warps executes the same instruction simultaneously, the different warps are scheduled by warp scheduler, when one warp is waiting for data loading the other warps begin to execute, this is called Single Instruction Multiple Thread (SIMT). This programming architecture is askin to Single Instruction Multiple Data (SIMD), however unlike the SIMD vectoring machines, SIMT can branch the behavior of each threads to achieve thread-level parallelism. The GPU use massive number of threads to hide the long latency of global memory instead of using large cache as CPU, in a word CPU cores are designed to minimize latency for each thread at a time, whereas GPUs are designed to handle a large number of concurrent, lightweight threads in order to maximize throughput.
\subsection{Mircroarchitecture of GPU} In Figure \ref{figure2} we use the GPU-GeForce GTX 760 as an example to show the microarchitecture of  GPU. The GPU works as a coprocessor system, consisting of Stream Multiprocessors (SM). The GeForce GTX 760 has 6 SM that perform the multiply-add arithmetic operation. Furthermore, each SM has the special functional units(SFU) that can perform more complex arithmetics such as trigonometric functions and reciprocal square root. The shared memory is on chip memory located at SM and has low latency and limited capacity. Global memory, which is off-chip memory of the GPU with long latency and large capacity. Another two types of the memory, called texture memory and constant memory, are located off-chip, but cached. Therefore the loading speed of the data located in these two type of memory space is much faster than that from global memory. Constant memory is read only and with latency approximately as small as the registers.  The variables invoked in kernel functions can be stored in either of the these four types of memories. Registers are allocated dynamically to threads.
\paragraph{}Because of the resource limitation of SM the total number of threads can be defined in one block is limited. On current CUDA supported GPUs, one block can contains up to 1024 threads, and all the blocks are executed on SM. Thus the total number of threads ($T_{total}$) we can make use of is $T_{total}=N_{threads}\times N_{threads/SM} $, $N_{SM}$ and $N_{threads/SM}$ denote the number of SM and number of threads per SM. The GeForce GTX 760 allows 2048 threads in one SM, thus the total number of threads that can work in parallel is $2048\times 6=12288$. In concrete applications, the programmer can allocate more than 12288 threads for a kernel, but the threads will work in serial, which will influence the program performance. 
\begin{figure}[htb]
\includegraphics[scale=0.5]{High_view_of_CUDA_GPU_microarchitecture.eps}
\caption{High View of CUDA CPU Microarchitecture}
\label{figure2}
\end{figure}   

@@

\section{GPU Based Acceleration of FCSD}\label{GPUFCSD}
In this section we present the parallel implementation of fixed complexity sphere decoder by GPU, first from (\ref{formula 1}), the maximum likelihood (ML) solution is given by:

@@

s_{ML}=arg\min_{s\in \mathbb{O}^{N_{t}}}||y-\mathbf{H}s||^{2} \label{formula 2}

@@

we consider the array $N_{t}=N_{r}$, performing QR decomposition to channel propagation matrix

@@

 \mathbf{H}=\mathbf{Q}\mathbf{R}  \label{QR}

@@

where $\mathbf{Q}\in \mathbf{C}^{N_{t}\times N_{t}}$ is the unitary matrix, $\mathbf{R}\in \mathbf{C}^{N_{t}\times N_{t}}$, is the upper triangular matrix, the basing on \cite{matrix theory}, the (\ref{formula 2}) can be changed to

@@

s_{ML}&=&arg\min_{s\in \mathbb{O}^{N_{t}}}||\mathbf{Q}^{H}y-\mathbf{Q}^{H}\mathbf{H}s||^{2}\\
&=& arg\min_{s\in \mathbb{O}^{N_{t}}}||\mathbf{Q}^{H}y-\mathbf{R}s||^{2} \label{formula 3}

@@

Consider the unconstrained estimation of $s$, $\hat{s}=\mathbf{G}y$, $\mathbf{G}=(\mathbf{H}^{H}\mathbf{H})^{-1}\mathbf{H}^{H}$ denotes the zero forcing equalizer, basing on \ref{QR}, we have  

>>>>

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[12pt,a4paper,final]{article}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
\usepackage[latin1]{inputenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
\usepackage{psfrag}
\usepackage{color}
\usepackage{epsfig}
\usepackage{appendix}
\usepackage{caption}
\usepackage{cite}
\usepackage{ifpdf}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{stfloats}
\usepackage{url}
\usepackage{fixltx2e}
\usepackage{setspace} 
\usepackage{diagbox}

@@

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Report}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
%\author{\IEEEauthorblockN{Tianpei Chen}
%\IEEEauthorblockA{ID: 25504946\\Department of Electrical and\\Computer Engineering\\
%McGill University\\
%Montreal, Quebec}}
\author{Tianpei Chen\\
 Department of Electrical and Computer Engineering\\}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
%\begin{abstract}

%\end{abstract}

@@

\section{System Model}\label{system}
% no \IEEEPARstart
%This demo file is intended to serve as a ``starter file''
%for IEEE conference papers produced under \LaTeX\ using
%IEEEtran.cls version 1.8 and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
%I wish you the best of success.
%\hfill mds
%\hfill December 27, 2012
Consider a MIMO system with $N_r$ receive antennas and $N_t$ transmit antennas, where $N_r$ is large, for simplification, we assume $N_r\to\infty$, the corresponding discrete time model is given:
\begin{equation}
y=\mathbf{H}s+n
\end{equation}
$s\in \mathbb{C}^{N_{t}\times 1}$ is the transmit symbol vector, $y\in \mathbb{C}^{N_{r}\times 1}$ is the receive symbol vector, $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ is the channel propagation matrix with independent identity distributed(i.i.d) Circularly Symmetric Complex Gaussian(CSCG) zero mean elements of unit variance (Rayleigh fading), $n\in \mathbb{C}^{N_{r}\times 1}$ is the additive white Gaussian noise (AWGN) with zero mean elements of variance $\sigma^{2}$, $\varepsilon(ss^{H})=E_{s}\mathbf{I}_{Nt}$, $E_s$ denotes the average symbol energy, $\varepsilon$ denotes the expectation operation and $M$ denotes the constellation size.

@@

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}
\section{Diversity Maximization Antenna Selection Criterion}\label{channel partition}
First we consider a general case to illustrate how the diversity maximization antenna selection scheme works,
$N_{t}-N$ denotes the number of selected antennas and $\mathbf{H_{j}}\in \mathbb{C}^{N_{r}\times (N_{t}-N)}$ denotes the associated propagation matrix, $j\in {\displaystyle\biggl[1,{N_{t}\choose N}\biggr]}$, $s_{d}\in \mathbb{C}^{(Nt-N)\times 1}$ denotes the the transmit symbol vector, $y_{d}\in \mathbb{C}^{Nr\times 1}$ denotes the received symbol vector, $n\in \mathbb{C}^{Nr\times 1}$ denotes the Additive White Gaussian Noise (AWGN) with zero mean and variance $\sigma^{2}$ hence
y_{d}=\mathbf{H}_{j}s_{d}+n
matrix $\mathbf{G}_{j}\in \mathbb{C}^{(Nt-N)\times Nr}$ denote the linear equalizer(ZF/MMSE), the linear estimates after equalization $\hat{s_{d}}$ is expressed by:
\hat{s_{d}}=\mathbf{G}_{j}y_{d}=\mathbf{G}_{j}\mathbf{H}_{j}s_{d}+\mathbf{G}_{j}n\label{equation0}
From (\ref{equation0}) we can see, the equalizer colors the noise, the off-diagonal elements of $\mathbf{G}_{j}\mathbf{H}_{j}$ contributes to spatial inter symbol interference, vector $\mathbf{G}_{j}n$, contributes to additional noise, from (\ref{equation0}) we have:
[\hat{s_{d}}]_{m}=(gh)_{m}s_{d}+g_{m}n=g_{m}h_{m}[s_{d}]_{m}+\sum_{f\neq m}g_{m}h_{f}[s_{d}]_{f}+g_{m}n\label{equation01}
where $[\hat{s_{d}}]_{m}$ denotes the $m$th element of $\hat{s_{d}}$, $(gh)_{m}$ denotes the $m$th row of $\mathbf{G}\mathbf{H}_{j}$, $g_{m}$ denotes the $m$th row of $\mathbf{G}_{j}$ and $h_{m}$ denotes the $m$th column of $\mathbf{H}_{j}$, The post processing SINR of $m$th sub datastream and $j$th selected antenna subset is defined as\cite{3}:
\rho^{j}_{m}=\frac{E_{s}|g_{m}h_{m}|^{2}}{\sigma^{2}\parallel g_{m}\parallel^{2}+E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}}
$E_{s}$ and $\sigma$ are defined in section \ref{system}, $\sigma^{2}||g_{m}||^{2}$ is the energy of noise after equalization, which can be derived as follow:
\varepsilon[|g_{m}n|^{2}]= \varepsilon[(g_{m}n)(g_{m}n)^{H}]=\varepsilon[g_{m}nn^{H}g_{m}^{H}]=g_{m}\varepsilon[nn^{H}]g_{m}^{H}=g_{m}\sigma^{2}Ig_{m}^{H}
=\sigma^{2}\parallel g_{m} \parallel^{2}
as to $(gh)_{m}$, we have:
\varepsilon[(gh)_{m}s_{d}s_{d}^{H}(gh)_{m}^{H}]=(gh)_{m}\varepsilon[s_{d}s_{d}^{H}](gh)_{m}^{H}=(gh)_{m}E_{s}I(gh)_{m}^{H} =E_{s}||(gh)_{m}||^{2}\label{equation03}
In (\ref{equation03}), $E_{s}||(gh)_{m}||^{2}=E_{s}|g_{m}h_{m}|^{2}+E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}$, $E_{s}|g_{m}h_{m}|^{2}$ is the post processing energy of detected symbol, $E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}$ is the energy of spatial inter symbol interference. \\
Considering zero forcing algorithm first, the spatial equalizer $(\mathbf{G}_{j})_{ZF}=(\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}\mathbf{H}_{j}^{H}$, therefore from (\ref{equation0}), $\mathbf{G}_{j}\mathbf{H}_{j}=\mathbf{I}_{N_t-N}$ and
(\rho_{ZF}^{j})_{m}=\frac{E_{s}}{\sigma^{2}\parallel g_{m} \parallel^{2}}\label{equation05}
Let $\omega_{m}$ denote the $m$th row of $(\mathbf{H_{j}^{H}H_{j}})^{-1}$, we have
\parallel g_{m} \parallel^{2}=g_{m}g_{m}^{H}=\omega_{m}\mathbf{H}_{j}^{H}\mathbf{H}_{j}\omega_{m}^{H}\label{equation04}
In (\ref{equation04}) we have $(\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}\mathbf{H}_{j}^{H}\mathbf{H}_{j}=\mathbf{I}_{N_t-N}$, hence $\omega_{m}\mathbf{H_{j}}^{H}\mathbf{H_{j}}=e_{m}$, where $e_{m}$ deontes the row vector with $m$th element equals to 1, the others are all 0. Therefore (\ref{equation04}) can be written as:
\parallel g_{m} \parallel^{2}=((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}_{mm})^{H}
$((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}_{mm})^{H}$ denotes the $m$th diagonal element of $((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1})^{H}$. Because $\mathbf{H}_{j}^{H}\mathbf{H}_{j}$ is a Hermitian matrix, thus $((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1})^{H}=(\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}$\cite{7} and (\ref{equation05}) can be written as:
(\rho_{ZF}^{j})_{m}=\frac{E_{s}}{\sigma^{2}(\mathbf{H}_{j}^{H}\mathbf{H}_{j})_{mm}^{-1}}\label{equation02}
\end{equation}\\
Consider minimum mean square error (MMSE) detection, where $\mathbf{G}=\mathbf{G}_{MMSE}=(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}\mathbf{H}_{j}^{H}$, from \cite{5} \cite{6}, we have the expression of the post processing SINR of MMSE:
(\rho_{MMSE}^{j})_{m}=\frac{E_{s}}{\sigma^{2}(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}_{mm}}-1\label{equation3}
where $(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}_{mm}$ denotes the $m$th diagonal element of $(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$.
The criterion of the diversity maximization selection scheme is to choose the antenna subset which can maximize the worst post-processing SINR of the sub-datastream. The column elements of $\mathbf{H}_{j}$ are mutually independent Gaussian random variables, under favorable propagation condition, when $N_{r}\to \infty$ and the 2 norm of the column will become much larger than the inner product of two different columns. Hence, the columns of $\mathbf{H}_{j}$ become asymptotically orthogonal, a phenomenon that can be expressed as
\lim_{N_{r}\to \infty}\frac{(\mathbf{H}_{j}^{H}\mathbf{H}_{j})}{N_{r}}=\mathbf{I}_{N_t-N}
 Where $\mathbf{I}_{N_t-N}$ denotes an $(N_{t}-N)\times (N_{t}-N)$ identity matrix. Based on this property, for any $m$, $(\rho_{MMSE}^{j})_{m}$ and $(\rho_{ZF}^{j})_{m}$ in (\ref{equation02})(\ref{equation3}) are all the same, meaning that we can select arbitrarily the $N_{t}-N$ antenna subset.

@@

\section{Complexity Analysis of sel-MMSE and sel-MMSE-OSIC algorithm}\label{complexity}
\subsection{sel-MMSE}\label{sel-MMSE}
Define the cost of complex multiplication as 6 flops and complex addition as 2 flops, the sel-MMSE algorithm can be divided into 4 steps\cite{2}:\\
\textbf{1}. Antenna selection\\
Based on diversity maximization antenna selection criterion, divide the detection task into two parts:

@@

y=\mathbf{H}_{1}s_{1}+\mathbf{H}_{2}s_{2}+n\label{equation 4}

@@

Hence there are $N_{u}=$ $N_{t}\choose N$ possible antenna subsets, $\mathbf{H}_{1}\in \mathbb{C}^{N_{r}\times N}$ and $\mathbf{H_{2}}\in C^{N_{r}\times (N_{t}-N)}$.\\
\textbf{2}. As to $s_{1}\in \mathbb{C}^{N\times 1}$, we choose every possible symbol vector $s_{1}$ in the lattice and calculate:

@@

y^{k}=y-\mathbf{H}_{1}s_{1}^{k}   \label{equation41}

@@

$s_{1}^{k}$ denotes the $k$th possible $N\times 1$ symbol vector that is chosen in channel partition stage.
For ML detection:
x^{ML}=arg\min_{s}\parallel y-\mathbf{H}s \parallel^{2}\label{equation 5}
based on (\ref{equation 4}), (\ref{equation 5}) can be changed to
arg\min_{s_{1},s_{2}}\parallel y-\mathbf{H}_{1}s_{1}-\mathbf{H}_{2}s_{2} \parallel
At this stage sel-MMSE uses brute-force search, take every possible $s_{1}^{k}\in \mathbb{C}^{N\times 1}$ vector according to signal constellation into consideration, therefore there are $M^N$ possible $s_{1}^{k}$ \\
\textbf{3}. For every $s_{1}^{k}$ calculate $y_{k}=y-\mathbf{H}_{1}s_{1}^{k}$. There are $M^{N}$ possible $s_{1}^{k}$ as well as $y^{k}$. Using linear detection or OSIC to estimate $s_{2}^{k}$, the estimation is expressed as $\hat{s_{2}}^{k}$, then mapping $\hat{s_{2}}^{k}$ to the constellation to the nearest points, we use $s_{2}^{k}$ to represent it, every $s_{1}^{k}$ corresponds to one $s_{2}^{k}$, hence there are $M^{N}$ posssible $s_{2}^{k}$.\\
\textbf{4}. Find the solution $[s_{1}^{k},s_{2}^{k}]$ among $M^{N}$ candidates that has the minimum total Euclidean distance.\\
\hat{s}=arg\min_{k=1,2,\ldots,M^{N}}\parallel y^{k}-\mathbf{H}_{2}s_{2}^{k} \parallel^{2}\label{equation6}
According to section \ref{channel partition}, based on diversity maximization antenna selection criterion and channel hardening phenomenon we can choose an arbitrary antenna subset, and hence the cost of antenna selection is zero. 
The second step calculating $y^{k}=y-\mathbf{H}_{1}s_{1}^{k}$, including an matrix multiplication and vector substruction,
N_{r}N \label{cost11}
complex multiplications and
 N_{r}N  \label{cost21}
 complex substitution are required. Then calculate the MMSE estimation $\hat{s}_{2}^{k}$:
\hat{s}_{2}^{k}=[\mathbf{H}_{2}^{H}\mathbf{H}_{2}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H}_{2}^{H}y^{k}\label{equation7}
$\mathbf{H}_{2}\in \mathbb{C}^{N_{r}\times (N_{t}-N)}$, the calculation of weight matrix $\mathbf{G}_{MMSE}=[\mathbf{H}_{2}^{H}\mathbf{H}_{2}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H}_{2}^{H}$, because of the channel hardening phenomenon mentioned in section \ref{channel partition}, $\mathbf{G}_{MMSE}=[\mathbf{H}_{2}^{H}\mathbf{H}_{2}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H}_{2}^{H}=\frac{1}{N_{r}+\frac{\sigma^{2}}{E_{s}}}\mathbf{I}\mathbf{H}_{2}^{H}$, hence the calculation of $\mathbf{G}_{MMSE}$ is real-complex multiplication, which cost 2 flops, 1/3 of complex multiplication, hence get $\mathbf{G}_{MMSE}$ requires:
\frac{(N_{t}-N)N_{r}}{3}    \label{cost31}
\end{equation}\\
complex multiplication.\\
In the whole process $\mathbf{G}_{MMSE}$ only need to be calculated one time. Calculation of $\hat{s}_{2}^{k}=\mathbf{G}_{MMSE}y_{k}$ is the complex matrix multiplication which requires:
 (N_{t}-N)N_{r} \label{cost12}
 complex multiplications and 
 (N_{t}-N)(N_{r}-1) \label{cost22}
complex additions. Finally the calculation of MED in (\ref{equation6}) requires matrix multiplication, vector substruction which cost $ N_{r}(N_{t}-N)$ complex multiplications and $ N_{r}(N_{t}-N)$ complex additions and the calculation of 2-norm, which includes the calculation of absolute value and real number addition, calculation of absolute value cost 3 flops, which is $1/2$ of the cost of complex multiplication, real number addition cost 1 flop which is $1/2$ of the cost of complex addition. Hence the calculation of 2-norm costs $\frac{N_{r}}{2}$ complex multiplications and $\frac{N_{r}-1}{2}$ complex additions. Totally, the calculation of MED costs:
  N_{r}(N_{t}-N)+\frac{N_{r}}{2}\label{cost13}
 complex multiplications and
   N_{r}(N_{t}-N)+\frac{N_{r}-1}{2}\label{cost23}
complex additions are required. All the above are the cost of single candidate  $[s_{1}^{k},s_{2}^{k}]$, Totally as mentioned before there are $M^N$ possible solution candidates $[s_{1}^{k},s_{2}^{k}]$, therefore the sel-MMSE requires overall adding (\ref{cost11})(\ref{cost31})(\ref{cost12})(\ref{cost13}), please notice (\ref{cost31}) only need to be calculated one time in the whole process.
M^{N}(N_rN+(N_{t}-N)N_r+N_r(N_t-N)+\frac{N_r}{2})+\frac{(N_{t}-N)N_{r}}{3}=M^{N}(2N_{r}N_{t}-N_{r}N+\frac{N_r}{2})+\frac{(N_{t}-N)N_r}{3}\label{equation8}
complex multiplications and adding (\ref{cost21})(\ref{cost22})(\ref{cost23})
M^{N}(N_rN+(N_t-N)(N_r-1)+N_r(N_t-N)+\frac{N_r-1}{2})=M^{N}(2N_{r}N_{t}-N_{r}N-N_{t}+N+\frac{N_{r}-1}{2})\label{equation9}
complex additions, Based on (\ref{equation8})(\ref{equation9}) and with 6 flops for complex multiplication and 2 flops for complex addition, we have the total cost of sel-MMSE algorithm as

@@

f_{sel-MMSE}(N) &=& 6[M^{N}(2N_{r}N_{t}-N_{r}N+\frac{N_r}{2})+\frac{(N_{t}-N)N_r}{3}]\\  
&+&2M^{N}(2N_{r}N_{t}-N_{r}N-N_{t}+N+\frac{N_{r}-1}{2})\\
&=&M^{N}(16N_{r}N_{t}-8N_{r}N+4N_{r}-2N_{t}+2N-1)+2(N_{t}-N)N_{r}\quad flops  \label{equation10}

@@

According to \cite{2}, the complexity of sel-MMSE algorithm without making use of channel hardening phenomenon is 
f_{sel-MMSE_{O}}(N)&=&  N_{u}[4(N_t-N)^3+(N_t-N)^2(4N_r+1)+(N_t-N)(4N_r-6)+6]+N_r(N_t-N)\\
&&(8N_t-8N-2)+M^{N}(16N_{r}N_{t}-8N_{r}N+4N_{r}-1-2N_{t}+2N)\quad flops \label{equation52}
where $N_{u}={\displaystyle{N_{t}\choose N}}$. Table \ref{table_sel-MMSE} shows the complexity comparison between $f_{sel-MMSE}$and$f_{sel-MMSE_{O}}$, at channel partition stage they all use diversity maximization scheme in \cite{2} to determine $N$.
\begin{table}[t]
\begin{tabular}{|c|c|c|c|c|}
\diagbox{$M$}{$\frac{f_{sel-MMSE}}{f_{sel-MMSE_{O}}}$}{ $N_{r}\times N_{t}$}&$10\times 10$&$32\times 32$&$64\times 64$&$128\times 128$\\\hline

BPSK&0.0249&1.3746e-05&2.3406e-09&1.5672e-14\\\hline

QPSK&0.1683&4.3817e-04&5.9893e-07&3.2095e-11\\\hline

16QAM&0.9282&0.3098&0.0378&1.3460e-04\\\hline

64QAM&0.9988&0.9978&0.9996&0.9982\\
